---
title: "BigFlightHmwk"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nycflights13)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(hms)
```

To answer the question what time of day I should fly to avoid delays as much as possible we look at some flights data. I chose departure delay instead of arrival delay because I am more personally concerned about that than arrival. Based on my analysis I concluded that the only factor that considerably effects the departure time delay is the time of day in which you fly. I found no significant looking correlation with Airline, Origin or Temperature and their effects on average departure time.

The first thing I want to analyze is the time-of-day vs delay. I included the head of this data set in order to provide the reader with a better idea of the data I am working with. To begin this, I plotted a graph to show any visible relationship between these variables.

```{r , echo=FALSE, warning=FALSE}
head(flights %>% mutate(time = as_hms(ymd_hms(time_hour)))%>%
  group_by(time) %>% mutate(avg_delay=mean(dep_delay, na.rm = T)))
  
flights %>% mutate(time = as_hms(ymd_hms(time_hour)))%>%
  group_by(time) %>% mutate(avg_delay=mean(dep_delay, na.rm = T)) %>% 
  ggplot(mapping = aes(y=avg_delay, x =time)) + geom_point() + 
  ggtitle("Average Departure Delay by Time of Day") + ylab("Average Departure Delay (Min") + xlab("Time of Day")
```


From this graph we can see a relationship between Time of Day and Average Departure Delay in minutes. From the graph we can conclude that if you want to avoid delays as much as possible you should fly before noon. The average Departure delay is bout -5 minutes, meaning leaving early, if you fly before 8am, close to an on time departure if you leave between 8am and noon. 

Next let’s look at another factor that may affect departure delays. Does Airline effect the average departure delay? I included the code for this section in order to display how I am piping variables. All of my sections are similar to this and vary by only a variable or two. Let’s plot some data and see:

```{r, warning=FALSE}
#AVG DELAY BY CARRIER
flights %>% mutate(time = as_hms(ymd_hms(time_hour)))%>%
  group_by(time, carrier) %>% mutate(avg_delay=mean(dep_delay, na.rm = T)) %>% 
  ggplot(mapping = aes(y=avg_delay, x = carrier, color = time) ) + geom_point() + 
  ggtitle("Average Departure Delay by Time of Day and Airline") + ylab("Average Departure Delay") + xlab("Airline")
```

This data doesn't display a super clear relationship, we can infer some things about it though. Most airlines appear to be consistent with some departure delay minutes, most lying in between 0 to 25 minutes. 

Now we look at if Airport perhaps effects departure delay:

```{r, echo=FALSE, warning=FALSE}
#AVG DELAY BY AIRPORT
flights %>% mutate(time = as_hms(ymd_hms(time_hour)))%>%
  group_by(time, origin) %>% mutate(avg_delay=mean(dep_delay, na.rm = T)) %>% 
  ggplot(mapping = aes(y=avg_delay, x = origin, color = time) ) + geom_point() + 
  ggtitle("Average Departure Delay by Time of Day and Airport of Origin") + ylab("Average Departure Delay") + xlab("Airport")
```

From this graph I can see no clear pattern, I conclude that airport does not have a large effect on departure delay. 

Now I explore if temperature had is a factor of departure delay: 

```{r, echo=FALSE, warning=FALSE}
#AVG DELAY BY TEMP
flights %>%
  left_join(weather) %>%
  mutate(time = as_hms(ymd_hms(time_hour)))%>%
  group_by(time, temp) %>% mutate(avg_delay=mean(dep_delay, na.rm = T)) %>% 
  ggplot(mapping = aes(y=avg_delay, x = temp, color = time) ) + geom_point() + 
  ggtitle("Average Departure Delay by Time of Day and Weather") + ylab("Average Departure Delay") + xlab("Tempurature")
```

From this graph I see no clear pattern to indicate significant effect, though it is interesting to look at. Visually it looks like the flights from 25 to 50 degrees looks more compact than the flights between 75 to 100 degrees.
